---
title: "RWork-sheet_jalando-on#4b"
author: "Ralyn Queen Jalando-on"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**USING FOR LOOP FUNCTION**

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r}
matrixVA <- matrix(0, nrow = 5, ncol = 5)
vectorA <- c(1, 2, 3, 4, 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrixVA[i, j] <- abs(i - j)
  }
}

print(matrixVA)
  
```

2. Print the string "*" using for() function. The output should be the same as shown
in Figure
```{r}
rows <- 5

for (i in 1:rows) {
  cat(rep("''*''", i), "\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}

```
**USING BASIC GRAPHICS (plot(),barplot(),pie(),hist())**

4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result
```{r}
library(readr)
data <- read_csv("/cloud/project/shoesizes.csv")
data
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}

df <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5,
                13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0,
             77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M",
             "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)

female_subset <- subset(df, Gender == "F")
male_subset <- subset(df, Gender == "M")

num_females <- nrow(female_subset)
num_males <- nrow(male_subset)

cat("Number of observations in Female:", num_females, "\n")

cat("Number of observations in Male:", num_males, "\n")
```
c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r}
gender_counts <- c(Female = 14, Male = 14) 

barplot(
  gender_counts,
  main = "Number of Male and Female in Household Data",  
  xlab = "Gender",                                        
  ylab = "Count",                                         
  col = c("#b4dd1e", "#9746f2"),                       
  legend.text = c("Female", "Male"),                    
  beside = TRUE                                           
)
```
5. The monthly income of Dela Cruz family was spent on the following:
a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r}
categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
values <- c(60, 10, 5, 25)

pie_chart <- pie(values, 
                  labels = paste(categories, "(", values, ")", sep = ""), 
                  col = c("#D3B99F", "#EF8354", "#0F7173", "#DF3B57"),
                  main = "Monthly Income Distribution of Dela Cruz Family")

percentages <- round(values / sum(values) * 100, 1)
text(0, 5, paste(percentages, "%"), cex = 1.2, pos = 3)

```

6. Use the iris dataset.
data(iris)
a. Check for the structure of the dataset using the str() function. Describe what you
have seen in the output.
```{r}
data(iris)
str(iris)
```
The iris dataset is a data frame with 150 obs and 5 columns. Four of these columns contain numeric measurements: Sepal.Length, Sepal.Width, Petal.Length, and Petal.Width, which describe the physical features of iris flowers. The fifth column, Species, is a category that includes three types of iris: "setosa," "versicolor," and "virginica." 

b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r}
mean_values <- colMeans(iris[, 1:4])
mean_values
```
c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write
the R script and its result.
```{r}
data(iris)

species_counts <- table(iris$Species)

pie(species_counts,
    main = "Species Distribution in Iris Dataset",
    col = c("#6d98ba", "#D3B99F", "#C17767"),
    labels = paste(names(species_counts), "\n", species_counts))
legend("topright", legend = names(species_counts), fill = c("#6d98ba", "#D3B99F", "#C17767"))

```


d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.
```{r}
setosa_subset <- iris[iris$Species == "setosa", ]
versicolor_subset <- iris[iris$Species == "versicolor", ]
virginica_subset <- iris[iris$Species == "virginica", ]

last_six_setosa <- tail(setosa_subset, 6)
last_six_versicolor <- tail(versicolor_subset, 6)
last_six_virginica <- tail(virginica_subset, 6)

last_six_setosa
last_six_versicolor
last_six_virginica

```

e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,
     main = "Iris Dataset",
     sub = "Sepal Width and Length",
     xlab = "Sepal Length",
     ylab = "Sepal Width",
     pch = 19,
     col = iris$Species)

legend("topright", legend = levels(iris$Species),
       col = 1:3, pch = 19)

```

f. Interpret the result.
The pie chart shows how many of each iris species are in the dataset. Each slice represents the size of each group that determine which species is the most common. The subsets display the last six entries for each species, allowing us to look at their specific measurements, like sepal and petal sizes, and understand how they differ. The scatterplot illustrates the relationship between sepal length and width. If setosa clusters at lower measurements, it means this species tends to have smaller flowers, while if versicolor and virginica overlap, it suggests their sizes are similar, making it harder to tell them apart.

