---
title: "RWork-sheet_jalando-on#4b"
author: "Ralyn Queen Jalando-on"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**USING FOR LOOP FUNCTION**

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r}
matrixVA <- matrix(0, nrow = 5, ncol = 5)
vectorA <- c(1, 2, 3, 4, 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrixVA[i, j] <- abs(i - j)
  }
}

print(matrixVA)
  
```

2. Print the string "*" using for() function. The output should be the same as shown
in Figure
```{r}
rows <- 5

for (i in 1:rows) {
  cat(rep("''*''", i), "\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}

```
**USING BASIC GRAPHICS (plot(),barplot(),pie(),hist())**

4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result
```{r}
library(readr)
data <- read_csv("/cloud/project/shoesizes.csv")
data
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}

df <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5,
                13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0,
             77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M",
             "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)

female_subset <- subset(df, Gender == "F")
male_subset <- subset(df, Gender == "M")

num_females <- nrow(female_subset)
num_males <- nrow(male_subset)

cat("Number of observations in Female:", num_females, "\n")

cat("Number of observations in Male:", num_males, "\n")
```
c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r}
gender_counts <- c(Female = 14, Male = 14) 

barplot(
  gender_counts,
  main = "Number of Male and Female in Household Data",  
  xlab = "Gender",                                        
  ylab = "Count",                                         
  col = c("blue", "red"),                       
  legend.text = c("Female", "Male"),                    
  beside = TRUE                                           
)
```

