---
title: "RWorksheet_Jalando-on#4a"
author: "Ralyn Queen Jalando-on"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
table <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 
                13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0), 
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 
             77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0), 
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", 
             "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M"))
table
```
a. Describe the data.

The datasets contains shoesizes from 28 different individual that varies from 5.0 to 13.0. It also records their height that varies 59.0 to 77.0 inches and gender. 

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.
```{r}
maleData <- subset(table, Gender == "M", select = c(Shoe_size, Height))
maleData
femaleData <- subset(table, Gender == "F", select = c(Shoe_size, Height))
femaleData
```
c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
aveShoe <- mean(table$Shoe_size)
aveShoe
aveHeight <- mean(table$Height)
aveHeight
```
d. Is there a relationship between shoe size and height? Why?

Yes, there is a relationship between the shoe size and the height of each individual because their shoe sizes will correspond accordingly to their height.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
Consider data consisting of the names of months:
"March","April","January","November","January","September","October","September", "November","August","January","November","November","February","May","August", "July","December","August","August","September",
"November","February, "April")
```{r}
months_vector <- c("March","April","January","November","January","September","October","September", "November","August","January","November","November","February","May","August", "July","December","August","August","September",
"November","February", "April")

factor_months_vector <- factor(months_vector)
factor_months_vector
```
3. Then check the summary() of the months_vector and factor_months_vector. Inter-pret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector,)
summary(factor_months_vector)

```
Yes, they are both equally useful.

4. Create a vector and factor for the table below.
```{r}
direction <- c("East","West","North")
direction
frequency <- c(1, 4, 3)
frequency
factor_data <- factor(direction)
new_order_data <- factor(factor_data, levels = c("East","West","North"))
print(new_order_data)
```
5. Enter the data below in Excel with file name = import_march.csv
a. Import the excel file into the Environment Pane using read.table() function.
Write the code.
b. View the dataset. Write the R scripts and its result.
```{r}
library(readr)
import_march <- read_csv("/cloud/project/import_march.csv")
import_march
```

