---
title: "RWorksheet_Jalando-on#4c"
author: "Ralyn Queen Jalando-on"
date: "2024-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Use the dataset mpg
A data frame with 234 rows and 11 variables:
a. Show your solutions on how to import a csv file into the environment.
```{r}
mpgData <- read.csv("mpg.csv")
mpgData
```
b. Which variables from mpg dataset are categorical?

**Manufacturer, drv, cyl, model, trans, fl, class**

c. Which are continuous variables?

**dsipl, year, cyt, hwy**

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r}
library(dplyr)

manufacturersCount <- mpgData %>%
  group_by(manufacturer) %>%
  summarise(modelCount = n_distinct(model), .groups = 'drop') %>%
  arrange(desc(modelCount))

mostModelsManufacturer <- manufacturersCount %>% slice(1)

mostVariationCount <- mpgData %>%
  group_by(model) %>%
  summarise(variationCount = n(), .groups = 'drop') %>%
  arrange(desc(variationCount))
mostVariationModel <- mostVariationCount %>% slice(1)

mostModelsManufacturer 
mostVariationModel

```
a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
uniqueModels <- mpgData %>%
  group_by(manufacturer) %>%
  summarise(uniqueModels = list(unique(model))) %>%
  arrange(manufacturer)

uniqueModels
  
```

b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
modelCounts <- mpgData %>%
  group_by(manufacturer) %>%
  summarise(uniqueModelCount = n_distinct(model)) %>%
  arrange(desc(uniqueModelCount))

barplot(modelCounts$uniqueModelCount,
        names.arg = modelCounts$manufacturer,
        las = 2,
        col = "#FFE8E1",
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer",
        ylab = "Number of Unique Models")

library(ggplot2)

ggplot(modelCounts, aes(x = reorder(manufacturer, uniqueModelCount), y = uniqueModelCount)) + 
  geom_bar(stat = "identity", fill = "#FFE8E1") + 
  coord_flip() + labs(title = "Number of Unique Models by Manufacturer", 
                      x = "Manufacturer", 
                      y = "Number of Unique Models") + 
  theme_minimal()

```
2. Same dataset will be used. You are going to show the relationship of the model and the manufacturer.
a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
UniqueModels <- mpgData %>%
  group_by(manufacturer) %>%
  summarise(uniqueModels = list(unique(model))) %>%
  arrange(manufacturer)

uniqueModels
```

b. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}
modelCounts <- mpgData %>%
  group_by(manufacturer) %>%
  summarise(uniqueModelCount = n_distinct(model)) %>%
  arrange(desc(uniqueModelCount))

barplot(modelCounts$uniqueModelCount, 
        names.arg = modelCounts$manufacturer, 
        las = 2, 
        col = "#A10702", 
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models")


library(ggplot2)

ggplot(modelCounts, aes(x = reorder(manufacturer, uniqueModelCount), y = uniqueModelCount)) +
  geom_bar(stat = "identity", fill = "#A10702") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal()

```

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
library(ggplot2)
library(dplyr)
top_20_mpg <- mpgData %>%
  arrange(desc(cty)) %>%
  head(20)
ggplot(top_20_mpg, aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Top 20 Models by Year", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}
modelCounts <- mpg %>%
  group_by(model) %>%
  summarise(numberOfCars = n()) %>%
  arrange(desc(numberOfCars))

print(modelCounts)

```

a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.
```{r}
topModels <- mpgData %>%
  group_by(model) %>%
  summarise(numberOfCars = n()) %>%
  arrange(desc(numberOfCars)) %>%
  slice_head(n = 20)

ggplot(topModels, aes(x = reorder(model, numberOfCars), y = numberOfCars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +  
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Model") +
  theme_minimal() +
  scale_fill_manual(values = c("#ff1b8d", "#ffda00", "#1bb3ff", "#ff1b8d", "#ffda00",
                               "#1bb3ff", "#ff1b8d", "#ffda00", "#1bb3ff", "#ff1b8d",
                               "#ffda00", "#1bb3ff", "#ff1b8d", "#ffda00", "#1bb3ff",
                               "#ff1b8d", "#ffda00", "#1bb3ff", "#ff1b8d", "#ffda00"))
```

b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
library(viridis)
library(viridisLite)
topModels <- mpgData %>%
  group_by(model) %>%
  summarise(numberOfCars = n()) %>%
  arrange(desc(numberOfCars)) %>%
  slice_head(n = 20)

ggplot(topModels, aes(x = reorder(model, numberOfCars), y = numberOfCars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal() +
  scale_fill_manual(values = c("#ff0000", "#ff7f00", "#ffff00", "#00ff00", "#0000ff", 
                               "#8b00ff", "#ff0000", "#ff7f00", "#ffff00", "#00ff00",
                               "#0000ff", "#8b00ff","#ff0000", "#ff7f00", "#ffff00", 
                               "#00ff00", "#0000ff", "#8b00ff","#ff0000", "#ff7f00"))

```
5. Plot the relationship between cyl - number of cylinders and displ - engine displacement usinggeom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.
a. How would you describe its relationship? Show the codes and its result.
```{r}
ggplot(mpgData, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  theme_minimal() +
  scale_fill_viridis_c()

```

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?
```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point(size = 3, alpha = 0.7) + 
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (L)",
       y = "Highway MPG",
       color = "City MPG") +
  theme_minimal() +
  scale_color_viridis_c()
```

6. Import the traffic.csv onto your R environment.
```{r}
library(readr)
trafficData <- read_csv("/cloud/project/traffic.csv")
head(trafficData)
```

a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show your answer.
```{r}
numberOfObservations <- nrow(trafficData)
variables <- colnames(trafficData)
numberOfObservations
variables
```

b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}
library(dplyr)
junctions <- subset(trafficData, Junction == TRUE)
head(junctions)
```

c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
library(ggplot2)
ggplot(trafficData, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
  geom_line() +
  labs(title = "Vehicle Counts at Junctions Over Time",
       x = "Date and Time",
       y = "Number of Vehicles",
       color = "Junction") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

7. From alexa_file.xlsx, import it to your environment
a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.
```{r}
library(readxl)
alexaFile <- read_excel("/cloud/project/alexa_file.xlsx")
dimensions <- dim(alexaFile)
rows <- dimensions[1]
columns <- dimensions[2]

rows
columns 
```

b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.
```{r}
library(dplyr)
totalVariation <- alexaFile %>%
  group_by(variation) %>%
  summarise(total = n())
totalVariation
```

c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.
```{r}
library(ggplot2)

ggplot(totalVariation, aes(x = variation, y = total, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Counts for Each Variation", x = "Variation", y = "Total Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.
```{r}
alexaFile$verified_reviews <- as.numeric(alexaFile$verified_reviews)
reviewsByDate <- alexaFile %>%
  group_by(date) %>%
  summarise(totalVerifiedReviews = sum(verified_reviews, na.rm = TRUE))
library(ggplot2)
ggplot(reviewsByDate, aes(x = date, y = totalVerifiedReviews)) +
  geom_line(color = "#F55D3E") +
  labs(title = "Verified Reviews Over Time", x = "Date", y = "Number of Verified Reviews") +
  theme_minimal()
```

e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.
```{r}
library(ggplot2)
library(dplyr)

variationRatings <- alexaFile %>%
  group_by(variation) %>%
  summarise(ratingAverage = mean(rating, na.rm = TRUE))

ggplot(variationRatings, aes(x = variation, y = ratingAverage, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating by Variation", x = "Variation", y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

